---
title: "Analyzing gender Bias of Name Ethnicity Classification"
author: "Franziska Hafner"
date: "09/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(sf)
library(ggplot2)
library(dplyr)
```

# Analysis of Name Ethnicity Classification AI

### Creating dataframe for analysis
Here, the relevant csv files are read in, joined into one data frame, and an additional columns for whether or not the prediction was right (1 = right, 0 = false) is added.
```{r}
#21 and else analysis:
#true_nationality <- read.csv("/Users/franziskahafner/Desktop/paper/testing_AI/input/names_to_ethnicity_known_gender_21else.csv")
#predicted_nationality <- read.csv("/Users/franziskahafner/Desktop/paper/testing_AI/output/classified_known_gender_21else.csv")

#14 and else analysis:
true_nationality <- read.csv("/Users/franziskahafner/Desktop/paper/testing_AI/input/names_to_ethnicity_known_gender_and_age_14else.csv")
predicted_nationality <- read.csv("/Users/franziskahafner/Desktop/paper/testing_AI/output/classified_known_gender_and_age_14else.csv")

colnames(predicted_nationality) <- c("names", "predicted_nationality")
nationality_data <- inner_join(true_nationality, predicted_nationality)
cols <- c("nationality", "predicted_nationality")
#nationality_data$nationality <- as.factor(nationality_data$nationality)
#nationality_data$predicted_nationality <- factor(nationality_data$predicted_nationality, levels = levels(nationality_data$nationality))

nationality_data$right_prediction <- c()
for(i in seq(1,nrow(nationality_data))){
  nationality_data$nationality[i] <- gsub(" ", "", nationality_data$nationality[i])
  nationality_data$gender[i] <- gsub(" ", "", nationality_data$gender[i])
  nationality_data$predicted_nationality[i] <- gsub(" ", "", nationality_data$predicted_nationality[i])
  if (nationality_data$nationality[i] == nationality_data$predicted_nationality[i]){
    nationality_data$right_prediction[i] <- 1
  }
  else{
    nationality_data$right_prediction[i] <- 0
  }
}

nationality_data$gender <- as.factor(nationality_data$gender)
```

### New data frame storing information about right female/male predictions is created

```{r}
all_nationalities <- c(levels(as.factor(nationality_data$nationality)))
female_right_predictions <- rep(0,length(all_nationalities))
male_right_predictions <- rep(0,length(all_nationalities))

right_predictions_per_nat <- data.frame(all_nationalities, female_right_predictions, male_right_predictions)

for(i in seq(1,nrow(nationality_data))){
  if(as.character(nationality_data$gender[i])=="female"){
    current_nat <- nationality_data$nationality[i]
    new_value <- right_predictions_per_nat[right_predictions_per_nat$all_nationalities==current_nat, "female_right_predictions"] + nationality_data$right_prediction[i]
    right_predictions_per_nat[right_predictions_per_nat$all_nationalities==current_nat, "female_right_predictions"] <- new_value
  }
  else{
    current_nat <- nationality_data$nationality[i]
    new_value <- right_predictions_per_nat[right_predictions_per_nat$all_nationalities==current_nat, "male_right_predictions"] + nationality_data$right_prediction[i]
    right_predictions_per_nat[right_predictions_per_nat$all_nationalities==current_nat, "male_right_predictions"] <- new_value
  }
}
```
### Barplot about gender differences in accuracy

```{r}
title <- paste("Gender Differences in Accuracy ", length(all_nationalities)-1," Nationalities and Else")

p<-ggplot(data=right_predictions_per_nat) +
  geom_bar(stat='identity', aes(x=all_nationalities, y=male_right_predictions, fill = "#67c45f")) +
  geom_bar(stat='identity', aes(x=all_nationalities, y=female_right_predictions, fill = "#8e8bda"), width = 0.6) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(x="Nationalities", y="Right Predictions (out of 1000)", 
      title=title) +
       scale_fill_manual(name = "Gender", 
                    label = c("Male", "Female"), 
                    values = c("#67c45f", "#8e8bda"))
p
```

```{r}
all_nationalities <- c(levels(as.factor(nationality_data$nationality)))

TP_F <- rep(0,length(all_nationalities))
TP_M <- rep(0,length(all_nationalities))

FP_F <- rep(0,length(all_nationalities))
FP_M <- rep(0,length(all_nationalities))

FN_F <- rep(0,length(all_nationalities))
FN_M <- rep(0,length(all_nationalities))

TN_F <- rep(0,length(all_nationalities))
TN_M <- rep(0,length(all_nationalities))

sensitivity_F <- rep(0,length(all_nationalities))
sensitivity_M <- rep(0,length(all_nationalities))

specificity_F <- rep(0,length(all_nationalities))
specificity_M <- rep(0,length(all_nationalities))


Sensitivity_specificity_per_nat <- data.frame(all_nationalities, TP_F, TP_M, FP_F, FP_M, FN_F, FN_M, TN_F, TN_M, sensitivity_F, sensitivity_M, specificity_F, specificity_M)
#Sensitivity_specificity_per_nat[Sensitivity_specificity_per_nat$all_nationalities=="british", 8] <- 10

for(nationality in all_nationalities){
  for(i in seq(1,nrow(nationality_data))){
    #current nationality = nationality
    if(nationality_data$nationality[i]==nationality){
      #True Positive, if predicted = current nationality:
      if(nationality_data$predicted_nationality[i]==nationality){
        #true positive female is in column 2
        column <- 2
      }
      #False Negative, if predicted != current nationality:
      if(nationality_data$predicted_nationality[i]!=nationality){
        #false negative female is in column 6
        column <- 6
      }
    }
    #current nationality != nationality
    else{
      #False Positive, if predicted = current nationality:
      if(nationality_data$predicted_nationality[i]==nationality){
        #false positive female is in column 4
        column <- 4
      }
      #True Negative, if predicted != current nationality:
      if(nationality_data$predicted_nationality[i]!=nationality){
        #true negative female is in column 8
        column <- 8
      }
    }
    if(nationality_data$gender[i]=="male"){
      column <- column + 1
    }
    
    Sensitivity_specificity_per_nat[Sensitivity_specificity_per_nat$all_nationalities==nationality, column] <- Sensitivity_specificity_per_nat[Sensitivity_specificity_per_nat$all_nationalities==nationality, column]+1
  }
}

for(nationality in all_nationalities){
  #sensitivity female:
  TruePositiveFemale <- Sensitivity_specificity_per_nat[Sensitivity_specificity_per_nat$all_nationalities==nationality, "TP_F"]
  FalseNegativeFemale <- Sensitivity_specificity_per_nat[Sensitivity_specificity_per_nat$all_nationalities==nationality, "FN_F"]
  sensitivityFemale <- TruePositiveFemale/(TruePositiveFemale+FalseNegativeFemale)
  Sensitivity_specificity_per_nat[Sensitivity_specificity_per_nat$all_nationalities==nationality, "sensitivity_F"] <- sensitivityFemale
  
  #sensitivity male:
  TruePositiveMale <- Sensitivity_specificity_per_nat[Sensitivity_specificity_per_nat$all_nationalities==nationality, "TP_M"]
  FalseNegativeMale <- Sensitivity_specificity_per_nat[Sensitivity_specificity_per_nat$all_nationalities==nationality, "FN_M"]
  sensitivityMale <- TruePositiveMale/(TruePositiveMale+FalseNegativeMale)
  Sensitivity_specificity_per_nat[Sensitivity_specificity_per_nat$all_nationalities==nationality, "sensitivity_M"] <- sensitivityMale
  
  #specificity female:
  TrueNegativeFemale <- Sensitivity_specificity_per_nat[Sensitivity_specificity_per_nat$all_nationalities==nationality, "TN_F"]
  FalsePositiveFemale <- Sensitivity_specificity_per_nat[Sensitivity_specificity_per_nat$all_nationalities==nationality, "FP_F"]
  specificityFemale <- TrueNegativeFemale/(TrueNegativeFemale+FalsePositiveFemale)
  Sensitivity_specificity_per_nat[Sensitivity_specificity_per_nat$all_nationalities==nationality, "specificity_F"] <- specificityFemale
  
  #specificity male:
  TrueNegativeMale <- Sensitivity_specificity_per_nat[Sensitivity_specificity_per_nat$all_nationalities==nationality, "TN_M"]
  FalsePositiveMale <- Sensitivity_specificity_per_nat[Sensitivity_specificity_per_nat$all_nationalities==nationality, "FP_M"]
  specificityMale <- TrueNegativeMale/(TrueNegativeMale+FalsePositiveMale)
  Sensitivity_specificity_per_nat[Sensitivity_specificity_per_nat$all_nationalities==nationality, "specificity_M"] <- specificityMale
}
```

```{r}
Sensitivity_specificity_per_nat
```
### Proportion of True Positives

```{r}
title <- paste("Sensitivity ", length(all_nationalities)-1," Nationalities and Else")

p<-ggplot(data=Sensitivity_specificity_per_nat) +
  geom_bar(stat='identity', aes(x=all_nationalities, y=sensitivity_M, fill = "#67c45f")) +
  geom_bar(stat='identity', aes(x=all_nationalities, y=sensitivity_F, fill = "#8e8bda"), width = 0.6) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(x="Nationalities", y="Sensitivity", 
      title=title) +
       scale_fill_manual(name = "Gender", 
                    label = c("Male", "Female"), 
                    values = c("#67c45f", "#8e8bda"))
p
```
### Proportion of True Negatives

```{r}
title <- paste("Specificity ", length(all_nationalities)-1," Nationalities and Else")

p<-ggplot(data=Sensitivity_specificity_per_nat) +
  geom_bar(stat='identity', aes(x=all_nationalities, y=specificity_M, fill = "#67c45f")) +
  geom_bar(stat='identity', aes(x=all_nationalities, y=specificity_F, fill = "#8e8bda"), width = 0.6) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(x="Nationalities", y="Specificity", 
      title=title) +
       scale_fill_manual(name = "Gender", 
                    label = c("Male", "Female"), 
                    values = c("#67c45f", "#8e8bda"))
p
```

```{r}
nationality_data
```
```{r}
#table(nationality_data$year_of_birth, nationality_data$right_prediction)
years <- 1950:2000
right_prediction <- rep(0,length(years))
wrong_prediction <- rep(0,length(years))
accuracy <- rep(0,length(years))
age_differences <- data.frame(years, right_prediction, wrong_prediction, accuracy)
for(i in seq(1,nrow(nationality_data))){
  year <- nationality_data$year_of_birth[i]
  if(nationality_data$right_prediction[i]==1){
    age_differences[age_differences$years==year, "right_prediction"] <- age_differences[age_differences$years==year, "right_prediction"] + 1
  }
  else{
    age_differences[age_differences$years==year, "wrong_prediction"] <- age_differences[age_differences$years==year, "wrong_prediction"] + 1
  }
}

for (year in years){
  age_differences[age_differences$years==year, "accuracy"] <- age_differences[age_differences$years==year, "right_prediction"]/age_differences[age_differences$years==year, "wrong_prediction"]
}

```

```{r}
p <- ggplot(data=age_differences, aes(x=years, y=accuracy))+
  geom_line()+
  labs(x="Year of Birth", y="Proportion of Right/Wrong Prediction", 
      title="Proportion of Right/Wrong Prediction by Age")
p
  
```





